Sensitivity Analysis

### 2024-05-28 Sections/Model.tex
28.05

	\subsection{Governing equations} \label{CH:Governing_equations_chapter}
	Following the work of \citet{Anderson1995}, the governing equations for quasi-one-dimensional were derived. Quasi-one-dimensional flow refers to a fluid flow scenario assuming that flow properties are uniformly distributed across any cross-section. This simplification is typically applied when the flow channel's cross-sectional area changes, such as through irregular shapes or partial fillings of an extractor. According to this assumption, velocity and other flow properties change solely in the flow direction.

@@

	The set of quasi-one-dimensional governing equations in Cartesian coordinates is described by Equations \ref{EQ: CompressibleEuler_1} - \ref{EQ: CompressibleEuler_3}.

@@

	where $\rho_f$ is the density of the fluid, $A_f$ is the function which describe change of the cross-section, $v$ is the velocity, $P$ is the total pressure, $e$ is the internal energy of the fluid, $t$ is time and $z$ is the spacial direction.

@@

	The previously derived quasi-one-dimensional continuity equation (Equation \ref{EQ: CompressibleEuler_1}) is redefined by incorporating a function $A_f = A\phi$. This modification distinguishes constant and varying terms, where the varying term accounts for the changes in the cross-sectional area available for the fluid. Equation \ref{EQ: Continuity_differential} shows the modified continuity equation:

@@

	To simplify the system's dynamics, it is assumed that $F$ is a control variable and affects the whole system instantaneously (due to $\nabla \cdot u = 0$), which allows finding the velocity profile that satisfies mass continuity based on $F$, $\phi$, and $\rho_f$.

@@

	Similarly, the superficial velocity might be introduced.

@@

	The fluid density $\rho_f$ can be obtained from an equation of state ({\color{red}article 1}) if temperature and thermodynamic pressure are known along $z$. The fluid density variation may occur due to pressure or inlet temperature changes. In the non-isothermal case, $\rho_f$ in Equations \ref{EQ:Velocity_1} and \ref{EQ:Velocity_2} is considered an average fluid density along the extraction column.

@@

	Equation \ref{Model_fluid} describes the movement of the solute in the system, which is constrained to the axial direction due to the quasi-one-dimensional assumption. Given that the solute concentration in the solvent is negligible, the fluid phase is described as pseudo-homogeneous, with properties identical to the solvent itself. It is also assumed that the thermodynamic pressure remains constant throughout the device. The analysis further simplifies the flow dynamics by disregarding the boundary layer near the extractor's inner wall. This leads to a uniform velocity profile across any cross-section perpendicular to the axial direction. Thus, the mass balance equation includes convection, diffusion, and kinetic terms representing the fluid phase behaviour.

@@

	$c_f$ represents the solute's concentration in the fluid phase, $r_e$ is a mass transfer kinetic term, and $D^M_e$ is the axial diffusion coefficient.

@@

	As given by Equation \ref{Model_solid}, the solid phase is considered stationary, without convection and diffusion terms in the mass balance equation. Therefore, the only significant term in this equation is the kinetic term of Equation \ref{Model_kinetic_basic}, which connects the solid and fluid phases. For simplicity, the extract is represented by a single pseudo-component.

@@

			As the solvent flows through the bed, CO$_2$ molecules diffuse into the pores and adsorb on the particle surface to form an external fluid film around the solid particles due to the solvent-solid matrix interactions. The dissolved solute diffuses from the particle's core through the solid-fluid interface, the pore, and the film into the bulk. Figure \ref{fig: SFE_Mechanism} shows the mass transfer mechanism, where the mean solute concentration in the solid phase is denoted as $c_s$, and the equilibrium concentrations at the solid-fluid interface are denoted as $c_s^*$ and $c_p^*$, respectively, for solid and fluid phases. The concentration of the solutes in the fluid phase in the centre of the pore is denoted as $c_p$. As the solute diffuses through the pore, its concentration changes and reaches $c_{pf}$ at the pore's opening. Then, the solute diffuses through the film around the particle and reaches bulk concentration $c_f$. The two-film theory describes the solid-fluid interface inside the pore. The overall mass transfer coefficient can be determined from the relationship between the solute concentration in one phase and its equilibrium concentration.

			\begin{figure}[h!]
				\centering
				\includegraphics[trim = 45cm 0cm 60cm 20cm,clip,width=0.85\columnwidth]{Figures/SFE_PFD.drawio.png}	
				\caption{The mass transfer mechanism}
				\label{fig: SFE_Mechanism}
			\end{figure}

			\citet{Bulley1984} suggests a process where the driving force for extraction is given by the difference between the concentration of the solute in bulk, $c_f$, and in the centre of the pore, $c_p^*$. The concentration $c_p^*$ is in equilibrium with $c_s$ according to an equilibrium relationship. The rate of extraction is thus $r_e\left(c_f - c^*_p(c_s)\right)$. On the other hand, \citet{Reverchon1996} proposes a driving force given by the difference between $c_s$ and $c_p^*$. Concentration $c_p^*$ is determined by an equilibrium relationship with $c_f$ and the extraction rate given by Equation \ref{Model_kinetic_basic}

@@

			where $\mu$ is sphericity, $l$ a characteristic dimension of particles and can be defined as $l = r/3$, $r$ is the mean particle radius, $\rho_s$ is the solid density, $D_i$ corresponds to the overall diffusion coefficient and $c_P^*$ is a concentration at the solid-fluid interface (which according to the internal resistance model is supposed to be at equilibrium with the fluid phase).

			According to \citet{Bulley1984}, a linear equilibrium relationship (Equation \ref{Linear_equilibirum}) can be used to find an equilibrium concentration of the solute in the fluid phase $c_f^*$ based on the concentration of the solute in the solid phase $c_s$.

@@

			The volumetric partition coefficient $k_p$ acts as an equilibrium constant between the solute concentration in one phase and the corresponding equilibrium concentration at the solid-fluid interphase. According to \citet{Spiro2007}, $k_p$ can be expressed through the mass partition coefficient $k_m$.

@@

			The process yield is calculated according to Equation \ref{Model_measurment_1} as presented by \citet{Sovova1994a}. The measurement equation evaluates the solute's mass at the extraction unit's outlet and sums it up. The integral form of the measurement (Equation \ref{Model_measurment_1}) can be transformed into the differential form (Equation \ref{Model_measurment_2}) and augmented with the process model.

@@

			The method of lines is used to transform the process model equations into a set of ODEs denoted as $G(x;\Theta)$. The backward finite difference is used to approximate the first-order derivative, while the central difference scheme is used to approximate the second-order derivative $z$-direction. The length of the fixed bed is divided into $N_z$ equally distributed points in $z$-direction. The state-space model after the discretization is denoted as $x$ and defines as follow:

@@

			For a derivative to be conservative, it must form a telescoping series. In other words, after the addition of all terms coming from the discretization over a grid, only the boundary terms should remain, and the artificial interior points should cancel out. The discretization is applied to the conservative form of the model to ensure mass conservation.

>>>>

	\subsection{Governing equations} \label{CH:Governing_equations_chapter}
	Following the work of \citet{Anderson1995}, the governing equations for a quasi-one-dimensional flow were derived. A quasi-one-dimensional flow refers to a fluid flow scenario assuming that the flow properties are uniformly distributed across any cross-section. This simplification is typically applied when the flow channel's cross-sectional area changes, such as through irregular shapes or partial filling of an extractor. According to this assumption, velocity and other flow properties change solely in the flow direction.

@@

	The set of quasi-one-dimensional governing equations in Cartesian coordinates is described by Equations \ref{EQ: CompressibleEuler_1} - \ref{EQ: CompressibleEuler_3}:

@@

	where $\rho_f$ is the density of the fluid, $A_f$ is the function which describes a change in the cross-section, $v$ is the velocity, $P$ is the total pressure, $e$ is the internal energy of the fluid, $t$ is time and $z$ is the spatial direction.

@@

	The previously derived quasi-one-dimensional continuity equation (Equation \ref{EQ: CompressibleEuler_1}) is redefined by incorporating the function $A_f = A\phi$. This modification distinguishes constant and varying terms, where the varying term accounts for changes in the cross-sectional area available for the fluid. Equation \ref{EQ: Continuity_differential} shows the modified continuity equation:

@@

	To simplify the system dynamics, it is assumed that $F$ is a control variable and affects the whole system instantaneously (due to $\nabla \cdot u = 0$), which allows finding the velocity profile that satisfies mass continuity based on $F$, $\phi$ and $\rho_f$:

@@

	Similarly, superficial velocity may be introduced:

@@

	The fluid density $\rho_f$ can be obtained from an equation of state (Appendix \ref{subsubsec: Equation of state}) if the temperature and thermodynamic pressure are known along $z$. Variation in fluid density may occur due to pressure or inlet temperature changes. In a non-isothermal case, in Equations \ref{EQ:Velocity_1} and \ref{EQ:Velocity_2} $\rho_f$ is considered the average fluid density along the extraction column.

@@

	Equation \ref{Model_fluid} describes the movement of the solute in the system, which is constrained to the axial direction due to the quasi-one-dimensional assumption. Given that the solute concentration in the solvent is negligible, the fluid phase is described as pseudo-homogeneous, with properties identical to those of the solvent itself. It is also assumed that the thermodynamic pressure remains constant throughout the device. The analysis further simplifies the flow dynamics by disregarding the boundary layer near the extractor's inner wall. This leads to a uniform velocity profile across any cross-section perpendicular to the axial direction. Thus, the mass balance equation includes convection, diffusion and kinetic terms representing the fluid phase behaviour:

@@

	where $c_f$ represents the solute concentration in the fluid phase, $r_e$ is the mass transfer kinetic term and $D^M_e$ is the axial diffusion coefficient.

@@

	As given by Equation \ref{Model_solid}, the solid phase is considered stationary, without convection and diffusion terms in the mass balance equation. Therefore, the only significant term in this equation is the kinetic term of Equation \ref{Model_kinetic_basic}, which connects the solid and fluid phases. For simplicity, the extract is represented by a single pseudo-component:

@@

			As the solvent flows through the bed, CO$_2$ molecules diffuse into the pores and adsorb on the particle surface to form an external fluid film around the solid particles due to the solvent-solid matrix interactions. The dissolved solute diffuses from the particle's core through the solid-fluid interface, the pore and the film into the bulk. Figure \ref{fig: SFE_Mechanism} shows the mass transfer mechanism, where the mean solute concentration in the solid phase is denoted as $c_s$, and the equilibrium concentrations at the solid-fluid interface are denoted as $c_s^*$ and $c_p^*$ for the solid and fluid phases, respectively. The concentration of the solutes in the fluid phase in the centre of the pore is denoted as $c_p$. As the solute diffuses through the pore, its concentration changes and reaches $c_{pf}$ at the pore opening. Then, the solute diffuses through the film around the particle and reaches bulk concentration $c_f$. The two-film theory describes the solid-fluid interface inside the pore. The overall mass transfer coefficient can be determined from the relationship between the solute concentration in one phase and its equilibrium concentration.

			\begin{figure}[h!]
				\centering
				\includegraphics[trim = 45cm 0cm 60cm 20cm,clip,width=0.85\columnwidth]{Figures/SFE_PFD.drawio.png}	
				\caption{Mass transfer mechanism.}
				\label{fig: SFE_Mechanism}
			\end{figure}

			\citet{Bulley1984} suggest a process where the driving force for extraction is given by the difference between the concentration of the solute in the bulk, $c_f$, and in the centre of the pore, $c_p^*$. The concentration $c_p^*$ is in equilibrium with $c_s$ according to the equilibrium relationship. The rate of extraction is thus $r_e\left(c_f - c^*_p(c_s)\right)$. In contrast, \citet{Reverchon1996} proposes a driving force given by the difference between $c_s$ and $c_p^*$. Concentration $c_p^*$ is determined by the equilibrium relationship with $c_f$ and the extraction rate given by Equation \ref{Model_kinetic_basic}:

@@

			where $\mu$ is sphericity, $l$ a characteristic dimension of particles that can be defined as $l = r/3$, $r$ is the mean particle radius, $\rho_s$ is the solid density, $D_i$ corresponds to the overall diffusion coefficient and $c_P^*$ is the concentration at the solid-fluid interface (which according to the internal resistance model is supposed to be at equilibrium with the fluid phase).

			According to \citet{Bulley1984}, a linear equilibrium relationship (Equation \ref{Linear_equilibirum}) can be used to find the equilibrium concentration of the solute in the fluid phase $c_f^*$ based on the concentration of the solute in the solid phase $c_s$:

@@

			The volumetric partition coefficient $k_p$ acts as an equilibrium constant between the solute concentration in one phase and the corresponding equilibrium concentration at the solid-fluid interphase. According to \citet{Spiro2007}, $k_p$ can be expressed through the mass partition coefficient $k_m$:

@@

			The process yield is calculated according to Equation \ref{Model_measurment_1} as presented by \citet{Sovova1994a}. The measurement equation evaluates the solute's mass at the extraction unit outlet and sums it up. The integral form of the measurement (Equation \ref{Model_measurment_1}) can be transformed into the differential form (Equation \ref{Model_measurment_2}) and augmented with the process model.

@@

			The method of lines is used to transform the process model equations into a set of ODEs denoted as $G(x;\Theta)$. The backward finite difference is used to approximate the first-order derivative, while the central difference scheme approximates the second-order derivative $z$ direction. The length of the fixed bed is divided into $N_z$, i.e. equally distributed points in the $z$ direction. The state-space model after discretization is denoted as $x$ and defined as follows:

@@

			For a derivative to be conservative, it must form a telescoping series. In other words, only the boundary terms should remain after adding all terms coming from the discretization over a grid, and the artificial interior points should be cancelled out. Discretization is applied to the conservative form of the model to ensure mass conservation.

### 2024-05-28 Sections/Saturation.tex
28.05

	Following the idea of the Broken-and-Intact Cell (BIC) model (\citet{Sovova2017}), the internal diffusion coefficient $D_i$ is considered to be a product of the reference value of $D_i^R$ and the exponential decay function $\gamma$, as given by Equation \ref{EQ: C_sat_function}.

@@

	where the ${\color{black}\Upsilon}$ describe the curvature of the decay function. Equation \ref{Model_kinetic} describes the final form of the kinetic term

@@

	The $\gamma$-function limits the solute's availability in the solid phase. Similarly to the BIC model, the solute is assumed to be contained in the cells, a part of which is open because the cell walls were broken by grinding, and the rest remains intact. The diffusion of the solute from a particle's core takes more time than the diffusion of the solute close to the outer surface. The same idea can be represented by the decaying internal diffusion coefficient, where the decreasing term is a function of the solute concentration in the solid.

	Alternatively, the decay function $\gamma$ can be interpreted by referring to the Shrinking Core model presented by \citet{Goto1996}, where the particle radius changes as the amount of solute in the solid phase decreases. As the particle size decreases due to dissolution, the diffusion path increases, which makes the diffusion slower and reduces the value of a diffusion coefficient. This analogy can be applied to the Equation \ref{EQ: C_sat_function} to justify the application of varying diffusion coefficient.

>>>>

	Following the idea of the Broken-and-Intact Cell (BIC) model (\citet{Sovova2017}), the internal diffusion coefficient $D_i$ is considered to be a product of the reference value of $D_i^R$ and the exponential decay function $\gamma$, as given by Equation \ref{EQ: C_sat_function}:

@@

	where  ${\color{black}\Upsilon}$ describes the curvature of the decay function. Equation \ref{Model_kinetic} describes the final form of the kinetic term:

@@

	The $\gamma$ function limits the solute's availability in the solid phase. Similarly to the BIC model, the solute is assumed to be contained in the cells, some of which are open because the cell walls were broken by grinding, with the rest remaining intact. The diffusion of the solute from a particle's core takes more time than the diffusion of the solute close to the outer surface. The same idea can be represented by the decaying internal diffusion coefficient, where the decreasing term is a function of the solute concentration in the solid.

	Alternatively, the decay function $\gamma$ can be interpreted by referring to the Shrinking Core model presented by \citet{Goto1996}, where the particle radius changes as the amount of solute in the solid phase decreases. As the particle size decreases due to dissolution, the diffusion path increases, which makes the diffusion slower and reduces the value of the diffusion coefficient. This analogy can be applied to Equation \ref{EQ: C_sat_function} to justify the application of a varying diffusion coefficient.
